### 2014-07-21 Studies/2012_Chicago/Report/Chicago_Fire.tex
Chicago: continued update to report write-up

\begin{table}
\centering
\captionof{table}{Fire Event Timeline}\label{tab:fire_info}
\begin{tabular}{l l}
\toprule[1.5pt]
Incident Time [hr:min:s] & Fire Behavior / Fireground Operation  \\
\midrule
\multirow{2}{*}{17:15:00} & \multirow{2}{*}{\parbox{10cm} {Reported working fire with heavy smoke coming from the front and rear of the structure's attic.}} \\ 
         & \\[.25cm] %blank rows exist for wrapping text. # of blanks is 1 less than number of rows
\multirow{2}{*}{17:16:00}    &  \multirow{2}{*}{\parbox{10cm} {Dispatch for a ``Still'' alarm $-$ two engine companies, two truck companies, and a battalion chief.}} \\
         & \\[.25cm]
\multirow{2}{*}{17:20:00}    & \multirow{2}{*}{\parbox{10cm} {Attic self ventilation - heavy smoke out of front attic window and heavy smoke and fire out of rear.}} \\
         & \\[.25cm] 
17:21:00 & E123 Capt. and FF on 2$^{nd}$ floor with 1 3/4-in hose line. \\[.25cm]
17:21:30 & 1$^{st}$ floor exterior porch door opened. \\[.25cm]
\multirow{2}{*}{17:23:00}    & \multirow{2}{*}{\parbox{10cm} {Heavy fire report in rear stairwell and enclosed porch. No response from E123: Capt. or FF.}} \\
         & \\[.25cm]
17:24:00    & Water applied to attic from the rear exterior position. \\[.25cm]
17:25:00    & SQ5 makes entry through front of structure. \\[.25cm]
17:27:00    & After hearing a Mayday call from SQ5, IC calls a Mayday. \\[.25cm]
17:29:00    & E123 Capt. carried out of structure. \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

\section{Geometry}
\label{geom}
The structure was a 2 1/2 story residential building with balloon frame construction and gable roof. Balloon frame construction is a method of framing where a continuous stud runs from the bottom horizontal member to the top plate. Intermediate floors are built directly off of the studs. One potential issue with balloon frame construction, is that without fire stops, there is void space for fire to potentially travel between the floors. The structure had a 17.15 m (56 ft) by 6.15 m (20 ft) footprint (cf. Fig.~(\ref{fig:simp_geom})). This results in a floor area of approximately 105 m$^2$ (1100 ft$^2$) per floor and a main living area of approximately 110 m$^2$ (2200 ft$^2$). Interior walls were constructed from nominal lumber with 12.7 mm (0.5 in) gypsum board over top. Additionally there was a half-story attic that was used to house the structure's hot water heater. There were three sets of stairs in the structure: one front stairwell that connected the first and second levels (Stair 1 in Fig.~(\ref{fig:simp_geom})), one rear stairwell that connected the first and second levels (Stair 2 in Fig.~(\ref{fig:simp_geom})) and a second rear set of stairs which provided access to the attic (Stair 3 in Fig.~(\ref{fig:simp_geom})). Entry to the first floor of the structure can be made either in the front or rear, see Fig.~(\ref{fig:simp_geom}). The fire, which is described in Section~\ref{fire}, started in the attic and spread to the rear porch. Discussion of the fire's impact on interior conditions can be found in Section~\ref{results}. Fig.~(\ref{fig:alpha_ex}) shows the post-incident exterior of the alpha side of the structure. Note that the majority of the damage is to the structure's attic.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=.65\textwidth]{../Figures/exterior_alpha}
\caption{Photograph showing the post-incident front (side alpha) exterior of the structure.}
\label{fig:alpha_ex}
\end{figure}

Figs.~(\ref{fig:first_floor} and \ref{fig:second_floor}) show the interior dimensions of the first and second floor, respectively. The dimensions were found from on-scene measurements using a tape measure. Measurements were made to the nearest 12.7 mm (0.5 in). Based on difficulty of making some of the measurements due to the condition of the post-fire structure, the uncertainty in the measurements is estimated to be on the order of $\pm$ 1 \%. This would result in measurements that could change by up $\pm$ 17 cm ($\approx$ 7 in). When considering the uncertainties in the initial fire development, ventilation (vent times and vent areas), and peak heat release rate among other factors, the uncertainty in the overall structure geometry is not a major factor.

%Fig.~(\ref{fig:smv_interior}) shows 3D Smokeview renderings of cutaways of the beta and delta sides of the structure interior.
%\begin{figure}[h!]
%\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}r}
%\includegraphics[width=.80\textwidth]{../Figures/interior_beta} \\
%\includegraphics[width=.80\textwidth]{../Figures/interior_delta} \\
%\end{tabular*}
%\caption{3D Smokeview rendering of structure interior looking from the charlie side to both the beta side (top) and to the delta side (bottom)}
%\label{fig:smv_interior}
%\end{figure}

@@

To estimate the fuel, fire size, and structure ventilation, on-scene videos and post-incident reports are used. Post-incident, the fire investigation concluded that the fire originated in the structure's attic~\cite{NIOSH:Bowyer}. Post-fire pictures revealed extensive damage to the attic and second floor rear porch (cf. Fig.~(\ref{fig:charlie_ex})). Based on these findings, the FDS model includes two source fire locations: an initial fire in the attic and a secondary fire on the second floor porch landing. The attic and rear porch consist of a fuel load that is mainly wood: the rear porch was enclosed with wood planks exteriorly covered by foam board and vinyl siding, the roof was asphalt shingles over wood planking, and the attic and porch had oriented strand board (OSB) floors.

\begin{figure}[h!]
\centering
\includegraphics[width=.65\textwidth]{../Figures/exterior_charlie}
\caption{Photograph showing the post-incident rear (side charlie) exterior of the structure.}
\label{fig:charlie_ex}
\end{figure}

If we represent wood with the following chemical formula, $\C\H_{1.7}\O_{0.74}\N_{0.002}$, and use carbon monoxide ($y_{\mathrm{CO}}=0.004 \; {\mathrm{kg_{CO}}/\mathrm{kg_{wood}}}$) and soot ($y_{\mathrm{C}}=0.015 \; {\mathrm{kg_{C}}/\mathrm{kg_{wood}}}$) yields from the SFPE Handbook~\cite{SFPE:Tewarson}, we can write a balanced chemical reaction for wood oxidation:

@@

The heat of combustion of wood used in this simulation is 16400 kJ/kg (7050 BTU/lb), based on data provided in the SFPE Handbook~\cite{SFPE:Tewarson}. The heat of combustion quantifies the energy release per unit mass of the fuel. The following text defines Eq.~\ref{eq:wood_comb} for an FDS input file with the desired fuel and reaction properties discussed above:

@@

To estimate the energy release per unit time and unit area (kW/m$^2$) of wood, we use work by Babrauskas and Grayson~\cite{babrauskas1990}. Experiments were conducted in a cone calorimeter to determine the 5 minute average of HRRPUA (kW/m$^2$) for several different types of wood over a range of radiant heat fluxes (kW/m$^2$). Here, we take a conservative approach and assume the burning wood in the structure had the lowest energy density of the tested samples~\cite{babrauskas1990}, similar to that of Douglas Fir plywood. Based on the range of heat flux exposures, the data was fit to determine HRRPUA ($y$) as a function of exposure ($x$):

\begin{equation}
y = ax+b
\label{eq:heat_flux}
\end{equation}

In Eq.~\ref{eq:heat_flux}, $a$ and $b$ are experimental constants. For Douglas Fir plywood these values were found to be 1.43 and 22 respectively. Typically radiative feedback from flaming combustion is between 20 kW/m$^2$ and 35 kW/m$^2$~\cite{tsaidrysdale}. If we estimate that the average radiation feedback to the pyrolyzing fuel is 20 kW/m$^2$ because of under-ventilated conditions within the structure, then following Eq.~\ref{eq:heat_flux} the average heat release per unit area (HRRPUA) for wood is 51 kW/m$^2$.

To calculate the total fire size, we need to determine the appropriate burnable area (exposed wood area) of the attic and second floor porch enclosure. Based on the floor plan of the structure, the underside of the roof is approximately 134 m$^2$ (1453 ft$^2$) with 7 m$^2$ (75 ft$^2$) walls on the Alpha and Charlie sides. Additionally there is 105 m$^2$ (1130 ft$^2$) of floor area in the attic. Examination of post incident images showed that the floor immediately over the back porch burned away. That area will be considered as part of the porch fire. Based on post-incident damage, we consider approximately 50 \% of the remaining floor area as fuel to the attic fire and 90 \% of the roof. To estimate the total attic fire size, we include the underside roof area (120 m$^2$), two end walls (14 m$^2$) and the floor area (43 m$^2$ (463 ft$^2$)). Using an HRRPUA of wood of 51 kW/m$^2$, results in a potential peak fire size of 9 MW. To estimate the porch fire size, we consider the the four walls of porch enclosure (42 m$^2$) and the floor and ceiling of the porch (24 m$^2$ (258 ft$^2$)). The total area is 64 m$^2$ (689 ft$^2$). Using the same HRRPUA of 51 kW/m$^2$, the potential peak fire size on the porch is approximately 3.3 MW. The total fire size for this structure is estimated to be at peak,12.3 MW. Since it is unknown when the fire started relative to notification of a fire or how fast the fire spread, several simplifications are made. First is that the attic fire starts at the beginning of the simulation and ramps up to 9 MW in 10 seconds. The porch fire starts 40 seconds into the simulation and ramps to 3.3 MW in 10 seconds. A ten second ramp to peak was used to prevent sharp velocity changes at the fire boundary condition in FDS. The second is that the sources of the gas phase wood fuel are of fixed area and at fixed locations (2 sources for attic, 1 source for porch). The following lines define the source fires for this FDS input file:

@@

\section{Materials}
\label{matl}
While the fire was simplified to constant area fuel sources (Section~\ref{fire}), defining the material properties of the walls and ceiling are still important from a heat transfer perspective. In this study, we define gypsum~\cite{Quintiere:2} for the finished walls and ceilings on the main two floors of the structure and wood~\cite{Incropera:1} for the unfinished portions of the attic and rear porch. The text provided below are the lines necessary to define these materials in this FDS input file.

@@

\section{Ventilation}
\label{Vents}
This simulation includes vents that account for the changes in ventilation due to a combination of fire department operations (opening doors) and fire acting on the structure (fire breaching walls, breaking windows, etc). There are two ventilation scenarios that are considered in this study. The first is the baseline simulation where the ventilation times and ventilation areas represent our best understanding of the incident. The alternative simulation does not include the firefighter opening the rear door to make entry to the structure and the subsequent beta side wall breach. The times of the ventilation changes are provided in Table \ref{tab:vents}. Structure leakage has been shown to be important when modeling enclosure fires~\cite{beal2009}. The open front door at the start of the simulation and the vents (opening size and open time) created during the simulation are significantly larger than the total effective area of leakage of structure of this type. Therefore, leakage would have a negligible impact on the fluid mechanics within the structure and is not included in this study.

\begin{table}
\centering
\captionof{table}{Simulation Ventilation Timelines}\label{tab:vents}
\begin{tabular}{ccccc}
\toprule[1.5pt]
Baseline Simulation [s] & Alternative Simulation [s] & Ventilation Area [m$^2$] & Type & Structure Side \\
\midrule
0    &  0   & 0.11 & Window      & Charlie \\
40    &  40   & 0.08 & Breach (x2) & Alpha and Charlie \\
45    &  45   & 0.12 & Breach      & Beta \\
55    &  55   & 0.12 & Breach      & Charlie \\     
130   &  N/A  & 1.41 & Door        & Charlie \\
135   &  N/A  & 0.2  & Breach      & Beta \\        
160   &  160  & 0.80 & Door        & Interior \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

For the simulation in this study, we can estimate a measure of how well the flow field is resolved by the non-dimensional expression $D^*/\dx$. Here $D^*$ is a characteristic fire diameter, $\dx$ is the nominal size of a mesh cell, and $\dQ$ is the total heat release rate of the fire:
\begin{equation}
D^* = \left(
     \frac{\dQ}{\rho_\infty \, c_p \, T_\infty \, \sqrt{g} }
     \right)^\frac{2}{5} 
\label{eq:mesh}
\end{equation}   
From the FDS User Guide~\cite{FDS_Users_Guide}, ``the characteristic fire diameter is related to the characteristic fire size via the
relation $Q^* = (D^*/D)^{5/2}$.'' Here, $D$ is the physical diameter of the fire as input to the model. Following Eq.~(\ref{eq:mesh}) and using 10 cm grid cells, the characteristic fire diameter to cell size ratio based on the 3.3 MW porch fire is 15.5 and for the 9 MW fire is 23. Based on validation work performed for the U.S. Nuclear Regulatory Commission, $D^*/\dx$ values ranged between 4 and 16~\cite{NUREG_1824}. The grid resolution used in this model falls within typical, engineering $D^*/\dx$ values. As a result, all of the input model geometry lengths, ventilation openings (holes, doors, and windows), and fire areas will snap to the nearest 10 cm. Fig.~(\ref{fig:geom_grid}) shows the 
alpha side (front) of the structure rendered in Smokeview with the 10 cm mesh.

\begin{figure}[h!]
\centering
\includegraphics[width=.60\textwidth]{../Figures/smv_exterior_grid}
\caption{Alpha side of the structure with 10 cm computation mesh.}
\label{fig:geom_grid}
\end{figure}

As discussed Section~\ref{geom}, the structure has 17.15 m (56 ft) by 6.15 m (20ft) footprint with a total height of 8.56 m (28 ft). To ensure adequately resolved fluid flow into and out of the structure the computational domain is extended beyond the volume of the structure. The total volume of the computational domain is 24 m (79 ft) by 10 m (33 ft) by 10 m (33 ft). At 10 cm (3.9 in) resolution this would result in 2.4 million computational cells. To make the simulation time more tractable, the domain was divided into 16 equal 150 thousand cell meshes to be processed in parallel. Fig.~(\ref{fig:mult_mesh}) shows the entire structure within the computational domain.

\begin{figure}[h!]
\centering
\includegraphics[width=.60\textwidth]{../Figures/smv_exterior_mesh}
\caption{Alpha-Delta side of the structure within computational domain. Each of the 16 meshes is 3 m (9.8 ft) by 10 m (33 ft) by 5 m (16.4 ft).}
\label{fig:mult_mesh}
\end{figure}

For the simulations performed for this study, Table ~\ref{tab:mod_param} includes the relevant model input parameters discussed in Section~\ref{model}.

\begin{table}
\centering
\captionof{table}{Relevant Modeling Input Parameters}\label{tab:mod_param}
\begin{tabular}{lll}
\toprule[1.5pt]
Parameter & Description & Reference  \\
\midrule
 Simulation Time           & 5 minutes                             & -- \\[.25cm]
 Grid Size                 & 10 cm                                 & Section~\ref{mesh} \\ [.25cm]
 Ambient Temperature       & 25 $^{\circ}$C                        & -- \\ [.1cm]
\multirow{4}{*}{Reaction: Wood}  & Formula: $\C\H_{1.7}\O_{0.74}\N_{0.002}$              & \multirow{4}{*}{Section~\ref{fire}} \\
                           & CO Yield: 0.004 \; ${\mathrm{kg_{CO}}/\mathrm{kg_{wood}}}$  & \\
                           & Soot Yield: 0.015 \; ${\mathrm{kg_{C}}/\mathrm{kg_{wood}}}$ & \\
                           & Heat of Combustion: 16400 kJ/kg                             & \\ [.25cm]
\multirow{2}{*}{Peak HRR}  & Attic Fire: 9 MW                                            & \multirow{2}{*}{Section~\ref{fire}}\\ 
                           & Porch Fire: 3.3 MW                                          & \\ [.25cm]                     
\multirow{3}{*}{Material: Wood}    & k: 0.12 \; \si{W/(m.K)})        & \multirow{3}{*}{Section~\ref{matl}} \\
                                   & $\rho$: 510 \; kg/m$^3$         & \\
                                   & c$_{p}$: 1.38 \; \si{kJ/(kg.K)} & \\[.25cm]
\multirow{3}{*}{Material: Gypsum}  & k: 0.17 \; \si{W/(m.K)}         &  \multirow{3}{*}{Section~\ref{matl}}\\ 
                                   & $\rho$: 930 \; kg/m$^3$        & \\ 
                                   & c$_{p}$: 1.09 \; \si{kJ/(kg.K)} & \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

\section{Heat Release Rate}
\label{HRR}
The heat release rate output data from FDS quantifies the simulated rate of energy release in the simulation. Here, we examine impact of ventilation on fire size. Fig.~(\ref{fig:hrr}) shows the HRR versus time based on the prescribed fires described in Section \ref{fire} and the FDS results for two venting scenarios. The two vertical lines represent when a firefighter opened the back door of the structure (t = 130 s) and when the second floor interior door was estimated to fail (t = 160 s).

@@

In Fig.~(\ref{fig:hrr}), the solid line is the baseline FDS simulation (this represents the on-scene conditions). The dashed line represents an alternative series events, specifically the rear door does not get opened and the subsequent beta side breach does not occur. The dash-dotted line is the prescribed FDS input HRR. We can see that within the first 10 seconds, the structure is under-ventilated as the modeled HRR is below the prescribed HRR. The HRR peaks at just under 9 MW (approximately 20 seconds into the simulation) before becoming oxygen limited. The HRR drops to 5 MW as the available oxygen inside the structure gets consumed. There is insufficient available oxygen in the structure to combust all of the fuel and combustion occurs locally at vents. Fig.~(\ref{fig:smv_ext_fire}) illustrates exterior flaming combustion through the failed Charlie side window and attic breach just after the second fire ignites at 40 seconds. After additional wall breaches at 45 and 55 seconds (see Table~\ref{tab:vents}), the abrupt change in ventilation causes the HRR spikes to 9 MW before again dropping to about 6 MW. The HRR then begins to increase towards a sustainable steady state value.

>>>>

\begin{table}
\centering
\captionof{table}{Fire Event Timeline}\label{tab:fire_info}
\begin{tabular}{l l}
\toprule[1.5pt]
Incident Time [hr:min:s] & Fire Behavior / Fireground Operation  \\
\midrule
\multirow{2}{*}{17:15:00} & \multirow{2}{*}{\parbox{10cm} {Reported working fire with heavy smoke coming from the front and rear of the structure's attic.}} \\ 
         & \\[.25cm] %blank rows exist for wrapping text. # of blanks is 1 less than number of rows
\multirow{2}{*}{17:16:00}    &  \multirow{2}{*}{\parbox{10cm} {Dispatch for a ``Still'' alarm $-$ two engine companies, two truck companies, and a battalion chief.}} \\
         & \\[.25cm]
\multirow{2}{*}{17:20:00}    & \multirow{2}{*}{\parbox{10cm} {Attic self ventilation - heavy smoke out of front attic window and heavy smoke and fire out of rear.}} \\
         & \\[.25cm] 
17:21:00 & E123 Capt. and FF on 2$^{nd}$ floor with 1 3/4 in hose line. \\[.25cm]
17:21:30 & 1$^{st}$ floor exterior porch door opened. \\[.25cm]
\multirow{2}{*}{17:23:00}    & \multirow{2}{*}{\parbox{10cm} {Heavy fire report in rear stairwell and enclosed porch. No response from E123: Capt. or FF.}} \\
         & \\[.25cm]
17:24:00    & Water applied to attic from the rear exterior position. \\[.25cm]
17:25:00    & SQ5 makes entry through front of structure. \\[.25cm]
17:27:00    & After hearing a Mayday call from SQ5, IC calls a Mayday. \\[.25cm]
17:29:00    & E123 Capt. carried out of structure. \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

\section{Geometry}
\label{geom}
The structure was a 2 1/2 story residential building with balloon frame construction and gable roof. \footnote{Balloon frame construction is a method of framing where a continuous stud runs from the bottom horizontal member to the top plate. Intermediate floors are built directly off of the studs. One potential issue with balloon frame construction, is that without fire stops, there is void space for fire to potentially travel between the floors.} The structure had a 17.15~m (56~ft) by 6.15~m (20~ft) footprint (cf. Fig.~(\ref{fig:simp_geom})). This results in a floor area of approximately 105~m$^2$ (1100~ft$^2$) per floor and a main living area of approximately 110~m$^2$ (2200~ft$^2$). Interior walls were constructed from nominal lumber with 12.7~mm (0.5~in) gypsum board over top~\cite{NIOSH:Bowyer}. Additionally there was a half-story attic that was used to house the structure's hot water heater. There were three sets of stairs in the structure: one front stairwell that connected the first and second levels (Stair 1 in Fig.~(\ref{fig:simp_geom})), one rear stairwell that connected the first and second levels (Stair 2 in Fig.~(\ref{fig:simp_geom})) and a second rear set of stairs which provided access to the attic (Stair 3 in Fig.~(\ref{fig:simp_geom})). Entry to the first floor of the structure can be made either in the front or rear, see Fig.~(\ref{fig:simp_geom}). The fire, which is described in Section~\ref{fire}, started in the attic and spread to the rear porch. Discussion of the fire's impact on interior conditions can be found in Section~\ref{results}. Fig.~(\ref{fig:alpha_ex}) shows the post-incident exterior of the front and rear of the structure. Note that the majority of the damage is to the structure's attic and second floor rear porch. Fully dimensioned drawings of the first and second floor interiors can be found in Figs.~(\ref{fig:first_floor} and \ref{fig:second_floor}), respectively.

@@

\begin{figure}[h!]
\centering
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}r}
   \scalebox{1.0}{ \includegraphics[width=3in]{../Figures/exterior_alpha} } &
   \scalebox{1.0}{ \includegraphics[width=2.67in]{../Figures/exterior_charlie} } \\
\end{tabular*}
\caption{Photograph showing the post-incident front (left) and rear (right) exterior of the structure.}
\label{fig:alpha_ex}
\end{figure}
%Fig.~(\ref{fig:smv_interior}) shows 3D Smokeview renderings of cutaways of the beta and delta sides of the structure interior.
%\begin{figure}[h!]
%\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}r}
%\includegraphics[width=.80\textwidth]{../Figures/interior_beta} \\
%\includegraphics[width=.80\textwidth]{../Figures/interior_delta} \\
%\end{tabular*}
%\caption{3D Smokeview rendering of structure interior looking from the charlie side to both the beta side (top) and to the delta side (bottom)}
%\label{fig:smv_interior}
%\end{figure}

@@

To estimate the fuel, fire size, and structure ventilation, on-scene videos and post-incident reports are used. Based on these findings, the FDS model includes two source fire locations: an initial fire in the attic and a secondary fire on the second floor porch landing. The attic and rear porch consisted of a fuel load that is mainly wood: the rear porch was enclosed with wood planks exteriorly covered by foam board and vinyl siding, the roof was asphalt shingles over wood planking, and the attic and porch had oriented strand board (OSB) floors.

Wood can be represented by the the following chemical formula, $\C\H_{1.7}\O_{0.74}\N_{0.002}$, with yields of carbon monoxide ($y_{\mathrm{CO}}=0.004 \; {\mathrm{kg_{CO}}/\mathrm{kg_{wood}}}$) and soot ($y_{\mathrm{C}}=0.015 \; {\mathrm{kg_{C}}/\mathrm{kg_{wood}}}$) found from the SFPE Handbook~\cite{SFPE:Tewarson}. A balanced chemical reaction for wood oxidation can written:

@@

The heat of combustion of wood used in this simulation is 16400 kJ/kg, based on data provided in the SFPE Handbook~\cite{SFPE:Tewarson}. The heat of combustion quantifies the energy release per unit mass of the fuel. The following text defines Eq.~\ref{eq:wood_comb} for an FDS input file with the desired fuel and reaction properties discussed above:

@@

To estimate the energy release per unit time and unit area (kW/m$^2$) of wood, we use work by Babrauskas and Grayson~\cite{babrauskas1990}. Experiments were conducted in a cone calorimeter \footnote{The cone calorimeter is an experimental apparatus used to gather data about ignition time, mass loss, combustion products, and heat release rate among other properties associated with burning small samples of materials.} to determine the 5 minute average of HRRPUA  \footnote{HRRPUA is the heat release rate per unit area.} (kW/m$^2$) for several different types of wood over a range of radiant heat fluxes (kW/m$^2$). Babrauskas and Grayson~\cite{babrauskas1990} suggest that the heat release rate per unit area for a variety of woods is approximately 50~kW/m$^2$.

To calculate the total fire size, the appropriate burnable area (exposed wood area) of the attic and second floor porch enclosure needs to be determined. Based on the floor plan of the structure, the underside of the roof is approximately 134~m$^2$ (1453~ft$^2$) with 7~m$^2$ (75~ft$^2$) walls on the front and rear sides. Additionally there is 105~m$^2$ (1130~ft$^2$) of floor area in the attic. Examination of post incident images showed that the floor immediately over the back porch burned away. That area will be considered as part of the porch fire. Based on post-incident damage, approximately 50 \% of the remaining floor area as fuel to the attic fire and 90 \% of the roof is con sidered. To estimate the total attic fire size, the underside roof area (120~m$^2$), two end walls (14~m$^2$) and the floor area (43~m$^2$ (463~ft$^2$)) are included. Using an HRRPUA of wood of 50~kW/m$^2$, results in a potential peak fire size of 9~MW. To estimate the porch fire size, the the four walls of porch enclosure (42~m$^2$) and the floor and ceiling of the porch (24~m$^2$ (258 ft$^2$)) are included. The total area is 64~m$^2$ (689~ft$^2$). Using the same HRRPUA of 50~kW/m$^2$, the potential peak fire size on the porch is approximately 3.3~MW. The total fire size for this structure is estimated to be at peak, 12.3~MW. Since it is unknown when the fire started relative to notification of a fire or how fast the fire spread, several simplifications are made. First is that the attic fire starts at the beginning of the simulation and ramps up to 9~MW in 10 seconds. The porch fire starts 40 seconds into the simulation and ramps to 3.3~MW in 10 seconds. A ten second ramp to peak was used to prevent sharp velocity changes at the fire boundary condition in FDS. The second is that the sources of the gas phase wood fuel are of fixed area and at fixed locations (2 sources for attic, 1 source for porch). The following lines define the source fires for this FDS input file:

@@

\section{Materials}
\label{matl}
While the fire was simplified to constant area fuel sources (Section~\ref{fire}), defining the material properties of the walls and ceiling are still important from a heat transfer perspective. In this study, gypsum~\cite{Drysdale:1} is used for the finished walls and ceilings on the main two floors of the structure and wood~\cite{Incropera:1} for the unfinished portions of the attic and rear porch. The text provided below are the lines necessary to define these materials in this FDS input file.

@@

\section{Ventilation}
\label{Vents}
The simulations in this study account for the changes in ventilation due to a combination of fire department operations (opening doors) and fire acting on the structure (fire breaching walls, breaking windows, etc). There are two ventilation scenarios that are considered in this study. The first is the baseline simulation where the ventilation times and ventilation areas represent our best understanding of the incident. The alternative simulation does not include the firefighter opening the rear door to make entry to the structure and the subsequent exterior left wall fire penetration. The times of the ventilation changes are provided in Table \ref{tab:vents}. Structure leakage has been shown to be important when modeling enclosure fires~\cite{beal2009}. The open front door at the start of the simulation and the vents (opening size and open time) created during the simulation are significantly larger than the total effective area of leakage of a structure of this type. Therefore, leakage would have a negligible impact on the fluid mechanics within the structure and is not included in this study.

\begin{table}
\centering
\captionof{table}{Simulation Ventilation Timelines}\label{tab:vents}
\begin{tabular}{ccccc}
\toprule[1.5pt]
Baseline Simulation [s] & Alternative Simulation [s] & Ventilation Area [m$^2$] & Type & Structure Side \\
\midrule
0    &  0   & 0.11 & Window      & Rear \\
40    &  40   & 0.08 & Breach (x2) & Front and Rear \\
45    &  45   & 0.12 & Breach      & Left \\
55    &  55   & 0.12 & Breach      & Rear \\     
130   &  N/A  & 1.41 & Door        & Rear \\
135   &  N/A  & 0.2  & Breach      & Left \\        
160   &  160  & 0.80 & Door        & Interior \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

For the simulation in this study, we can estimate a measure of how well the flow field is resolved by the non-dimensional expression $D^*/\dx$. Here $D^*$ is a characteristic fire diameter, $\dx$ is the nominal size of a mesh cell, and $\dQ$ is the total heat release rate of the fire:
\begin{equation}
D^* = \left(
     \frac{\dQ}{\rho_\infty \, c_p \, T_\infty \, \sqrt{g} }
     \right)^\frac{2}{5} 
\label{eq:mesh}
\end{equation}   
From the FDS User Guide~\cite{FDS_Users_Guide}, ``the characteristic fire diameter is related to the characteristic fire size via the
relation $Q^* = (D^*/D)^{5/2}$.'' Here, $D$ is the physical diameter of the fire as input to the model. Following Eq.~(\ref{eq:mesh}) and using 10 cm grid cells, the characteristic fire diameter to cell size ratio based on the 3.3 MW porch fire is 15.5 and for the 9 MW fire is 23. Based on validation work performed for the U.S. Nuclear Regulatory Commission, $D^*/\dx$ values ranged between 4 and 16~\cite{NUREG_1824}. The grid resolution used in this model falls within typical, engineering $D^*/\dx$ values. As a result, all of the input model geometry lengths, ventilation openings (holes, doors, and windows), and fire areas will snap to the nearest 10 cm. Fig.~(\ref{fig:geom_grid}) shows the 
front of the structure rendered in Smokeview with the 10 cm mesh.

\begin{figure}[h!]
\centering
\includegraphics[width=.60\textwidth]{../Figures/smv_exterior_grid}
\caption{Front side of the structure with 10 cm computation mesh.}
\label{fig:geom_grid}
\end{figure}

As discussed Section~\ref{geom}, the structure has 17.15~m (56~ft) by 6.15~m (20~ft) footprint with a total height of 8.56~m (28~ft). To ensure adequately resolved fluid flow into and out of the structure the computational domain is extended beyond the volume of the structure. The total volume of the computational domain is 24~m (79~ft) by 10~m (33~ft) by 10~m (33~ft). At 10~cm (3.9~in) resolution this would result in 2.4 million computational cells. To make the simulation time more tractable, the domain was divided into 16 equal 150 thousand cell meshes to be processed in parallel. Fig.~(\ref{fig:mult_mesh}) shows the entire structure within the computational domain.

\begin{figure}[h!]
\centering
\includegraphics[width=.60\textwidth]{../Figures/smv_exterior_mesh}
\caption{Front-right side of the structure within computational domain. Each of the 16 meshes is 3 m (9.8 ft) by 10 m (33 ft) by 5 m (16.4 ft).}
\label{fig:mult_mesh}
\end{figure}

For the simulations performed for this study, Table~\ref{tab:mod_param} includes the relevant model input parameters discussed in Section~\ref{model}.

\begin{table}
\centering
\captionof{table}{Relevant Modeling Input Parameters}\label{tab:mod_param}
\begin{tabular}{lll}
\toprule[1.5pt]
Parameter & Description & Reference  \\
\midrule
 Simulation Time           & 5 minutes                             & -- \\[.25cm]
 Grid Size                 & 10 cm                                 & Section~\ref{mesh} \\ [.25cm]
 Ambient Temperature       & 25 $^{\circ}$C                        & -- \\ [.1cm]
\multirow{4}{*}{Reaction: Wood}  & Formula: $\C\H_{1.7}\O_{0.74}\N_{0.002}$              & \multirow{4}{*}{Section~\ref{fire}} \\
                           & CO Yield: 0.004 \; ${\mathrm{kg_{CO}}/\mathrm{kg_{wood}}}$  & \\
                           & Soot Yield: 0.015 \; ${\mathrm{kg_{C}}/\mathrm{kg_{wood}}}$ & \\
                           & Heat of Combustion: 16400~kJ/kg                             & \\ [.25cm]
\multirow{2}{*}{Peak HRR}  & Attic Fire: 9~MW                                            & \multirow{2}{*}{Section~\ref{fire}}\\ 
                           & Porch Fire: 3.3~MW                                          & \\ [.25cm]                     
\multirow{3}{*}{Material: Wood}    & k: 0.12 \; \si{W/(m.K)})        & \multirow{3}{*}{Section~\ref{matl}} \\
                                   & $\rho$: 510 \; kg/m$^3$         & \\
                                   & c$_{p}$: 1.38 \; \si{kJ/(kg.K)} & \\[.25cm]
\multirow{3}{*}{Material: Gypsum}  & k: 0.17 \; \si{W/(m.K)}         &  \multirow{3}{*}{Section~\ref{matl}}\\ 
                                   & $\rho$: 930 \; kg/m$^3$        & \\ 
                                   & c$_{p}$: 1.09 \; \si{kJ/(kg.K)} & \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

\section{Heat Release Rate}
\label{HRR}
The heat release rate output data from FDS quantifies the simulated rate of energy release in the simulation. Fig.~(\ref{fig:hrr}) shows the HRR versus time based on the prescribed fires described in Section \ref{fire} and the FDS results for two venting scenarios. The two vertical lines represent when a firefighter opened the back door of the structure (t = 130 s) and when the second floor interior door was estimated to fail (t = 160 s). Due to a lack of information about when the fire ignited or how fast it grew, these parameters needed to be estimated.  While the growth rate shown in Fig.~(\ref{fig:hrr}) (10~MW in approximately 20 seconds) is atypically fast, analysis of the incipient stage of the fire growth was not considered in this study. The goal was to have a fire approach its estimated peak and become ventilation limited quickly to minimize computation time while aligning with the constraints of videos taken during the fire. Accuracy of heat transfer to the walls of the structure may be sacrificed by taking this approach, however all window, door, and wall failure were driven by the event timeline not modeling predicted failure temperatures.

@@

In Fig.~(\ref{fig:hrr}), the solid line is the baseline FDS simulation (this represents the on-scene conditions). The dashed line represents an alternative series events, specifically the rear door does not get opened and the subsequent beta side breach does not occur. The dash-dotted line is the prescribed FDS input HRR. We can see that within the first 10 seconds, the structure is under-ventilated as the modeled HRR is below the prescribed HRR. The HRR peaks at just under 9 MW before becoming oxygen limited. The HRR drops to 5~MW as the available oxygen inside the structure gets consumed. There is insufficient available oxygen in the structure to combust all of the fuel and combustion occurs locally at vents. Fig.~(\ref{fig:smv_ext_fire}) illustrates exterior flaming combustion through the failed Charlie side window and attic breach just after the second fire ignites at 40 seconds. After additional wall breaches at 45 and 55 seconds (see Table~\ref{tab:vents}), the abrupt change in ventilation causes the HRR spikes to 9~MW before again dropping to about 6~MW. The HRR then begins to increase towards a sustainable steady state value.

From Fig.~(\ref{fig:hrr}), we see that the HRR for both FDS simulations are identical until 130 seconds (HRR $\approx$ 7~MW). At 130 seconds the rear door is opened in the baseline simulation. At 135 seconds there is an additional beta side breach (Table~\ref{tab:vents}). The open door and breach act as additional pressure releases for the structure. Again, the abrupt change in ventilation causes the HRR to spike. Here we see a rise from 7~MW to 9.7~MW at 139 seconds. At the time of door failure (160 seconds), the HRR in both cases was approximately 9 MW as there was sufficient ventilation. While there might be some locally different burning characteristics between the baseline simulation and the alternate simulation, in both cases the interior conditions represented a significant hazard to firefighters inside the structure. Therefore, the analysis will focus on the results of the baseline simulation.
